---
title: "5xCreGraphTemplate"
author: "Rbass"
date: "2024-03-12"
output: html_document
---


In this basic analysis, i had a collection of images from 4 groups ("Cntl", "5xFAD", "Ntrk2-cKO" and "5xFAD;Ntrk2-cKO") in which the fluorescent intensity was measured using imageJ. ImageJ created a csv data frame that had all the file names (which were blinded with letters of the alphabet) and the corresponding fluorescent intensity measured. here, i de-blinded the data, manipulated the data frame with dlpyr, performed a t-test, and graphed the data using ggplot. 


```{r}
install.packages("ggpubr")
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
```

load data 
```{r}
subic <- read.csv("/Users/robertbass/Dropbox (UFL)/Xu_Lab/Data/5xFAD_astrocyte_tracing/6_month_5xCre_sagittal/5xCre_6_month_abeta_iba1/blinded images/_just 20x/subiculum/subiculumIntensity.csv")

somato <- read.csv("/Users/robertbass/Dropbox (UFL)/Xu_Lab/Data/5xFAD_astrocyte_tracing/6_month_5xCre_sagittal/5xCre_6_month_abeta_iba1/blinded images/_just 20x/somatomotor/somatomotorIntensity.csv")
```

generate averages of rows that are from the same sample. de-blind. generate p values.
```{r}
deblind <- function(df) {
  df$genotype <- "undefined"
  for (i in 1:nrow(df)) {
    if (df[i,]$letter == "C" | df[i,]$letter == "E" | df[i,]$letter == "F" | df[i,]$letter == "I") {
      df[i,]$genotype <- "CNTL"
    }
    else if (df[i,]$letter == "A" | df[i,]$letter == "B" | df[i,]$letter == "K" | df[i,]$letter == "N"
              | df[i,]$letter == "P" | df[i,]$letter == "Q") {
      df[i,]$genotype <- "5xFAD"
    }
    else if (df[i,]$letter == "D" | df[i,]$letter == "H" | df[i,]$letter == "J") {
      df[i,]$genotype <- "Ntrk2-cKO"
    }
    else if(df[i,]$letter == "G" | df[i,]$letter == "L" | df[i,]$letter == "M" | df[i,]$letter == "O") {
      df[i,]$genotype <- "5xFAD;Ntrk2-cKO"
    }
  }
  return(df)
}

statTest <- function(meanDF) {
  stats <- compare_means(y ~ genotype,
                data = meanDF, method = "t.test")
  return(stats)
}

#subiculum integrated density
subicIntDen <- subic %>% 
  separate(Label, c("letter", "section", "brain region", "objective", "channel"), sep = " ") %>% #separate file name into several columns
  group_by(letter) %>%
  summarise(y = mean(IntDen)) %>% #average rows that are from the same animal
  filter(!letter == "P") %>% #take out 5x 21, which doesn't have a beta plaques
  deblind() %>%
  arrange(genotype)
  
#somatomotor cortex
somatoIntDen <- somato %>% 
  separate(Label, c("letter", "section", "brain region", "objective", "channel"), sep = " ") %>%
  group_by(letter) %>%
  summarise(y = mean(IntDen)) %>%
  filter(!letter == "P") %>%
  deblind() %>%
  arrange(genotype)

subicIntDenStats <- statTest(subicIntDen)
somatoIntDenStats <- statTest(somatoIntDen)
```

```{r}
subicAreaStats <- statTest(subicArea)
somatoAreaStats <- statTest(somatoArea)
```





graph
```{r}
graphFunction <- function(df, stats, title) {
graph <- ggplot(data = df,
       # mapping = aes(x = genotype, y = y) #if you include aes here it will break the function which makes no sense
       ) +
  theme_light() +
  stat_summary( #bar graph
    # position = "dodge",
    mapping = aes(x = genotype, y = y, fill = genotype),
    fun = mean, fun.min = mean, fun.max = mean,
    geom = "bar", size = 0.1,
    width = 0.4
  ) +
  stat_summary( #error bars
    mapping = aes(x = genotype, y = y),
    position = position_dodge(width = 0.4), #if you just put "dodge" it will cling to the center
    fun = mean,
    fun.min = function(x) mean(x) - sd(x), #use sd() or std.error
    fun.max = function(x) mean(x) + sd(x),
    geom = "errorbar",
    size = 0.3, width = 0.2 #size is line thickness
  ) +
    geom_jitter(width = 0.1, mapping = aes(x = genotype, y = y, fill = genotype,
                                         # colour = brainSection#are you going to graph brain section?
                                         )
                ) +
  scale_x_discrete(
    limits = c("CNTL", "5xFAD", "Ntrk2-cKO", "5xFAD;Ntrk2-cKO"), #this sets order on x axis
    labels = c("CNTL", "5xFAD", "Ntrk2-cKO", "5xFAD;\nNtrk2-cKO")#this changes labels
    ) + 
  # scale_y_continuous(breaks = seq(0, 120, by = 25)) +
  # theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + #angle of the x variable text
  scale_fill_manual(
    values = c(#change colors of the bars
      "CNTL" = "grey",
      "5xFAD" = "dark green",
      "Ntrk2-cKO" = "red",
      "5xFAD;Ntrk2-cKO" = "blue"),
    breaks = c(  #this sets the order on the legend
      "CNTL", "5xFAD", "Ntrk2-cKO", "5xFAD;Ntrk2-cKO"),
    labels = c( #this labels the legend correctly
      "CNTL", "5xFAD", "Ntrk2-cKO", "5xFAD;Ntrk2-cKO"),
   ) +
  stat_pvalue_manual(stats, hide.ns = "p", label = "p.signif",
                     y.position = max(df$y)*1.1,
                     step.increase = 0.2,
                     size = 8 #this is for powerpoints
                     ) +
  labs(title = title,
       # x = "Genotype",
       # y = "Length (um)",
       # fill = "Genotype",
       ) +
  theme(plot.title = element_text(hjust = 0.5), #center the title
        legend.position = "none", #don't show the legend
        axis.title.x = element_blank() #get rid of x axis label "genotype"
        ) +
    #everything below this comment is for powerpoints. also, change the size of the p value number^
  scale_y_continuous(limits = c(0, max(df$y)*1.8)) +
  theme(plot.title = element_text(size = 25),
        axis.title.y = element_text(size = 18),
        axis.text.x = element_text(size = 14, color = "black"),
        # axis.text.x = element_text(size = 11, color = "black", angle = 45, vjust = 1, hjust = 1)) 
        # angle of the x variable text^^
        axis.text.y = element_text(size = 12, color = "black"),
        aspect.ratio = 8/10) #small number = short and thin
return(graph)
}

graphFunction(somatoIntDen, somatoIntDenStats, "Integrated Density, Somatomotor")
graphFunction(subicIntDen, subicIntDenStats, "Integrated Density, Subiculum")
```