---
title: "R Notebook"
output: html_notebook
---


In the following code, i analyzed this scRNAseq atlas of cells in the brain:
https://pubmed.ncbi.nlm.nih.gov/30096299/
I measured expression of BDNF and Ntrk2 in the frontal cortex and hippocampus of CNTL mice and exported that data as a PDF
```{r}
install.packages('/Users/xulab/Downloads/DropSeq.util_2.0.tar.gz', repos=NULL)

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("GEOquery")
install.packages("scCustomize")

#seurat 5.0.0 doesn't work with dge object. i know that 4.3.0 works
remotes::install_version("Seurat", version = "4.4.0")
# install.packages("Matrix", version='1.6-0')
# remotes::install_version("Matrix", version='1.6-0')
# 
# remotes::install_version("tidyverse", version='1.3.2')
# remotes::install_version("dplyr", version='1.1.2')
```


```{r}
library("dplyr")
library("tidyverse")
library("Seurat")
library("ggplot2")
library("cowplot")
library("ggpubr")
library("DropSeq.util")
library("scCustomize")
# library("GEOquery")
# install.packages("Rcpp", dependencies = TRUE)

available.packages("seurat")
```


# script to perform standard workflow steps to analyze single cell RNA-Seq data

create the seurat object (raw,non-normalized data)
```{r}
dge <- loadSparseDge("/Users/xulab/Documents/Rob/changran analysis/Molecular Diversity and Specialization/F_GRCm38.81.P60Cortex_noRep5_FRONTALonly.raw.dge.txt.gz") 


seu <- CreateSeuratObject(counts = dge, project = "test", min.cells = 3, min.features = 200)
# beep(1)
# view(seu@meta.data)
# seu <- seu2
```




```{r}
# there are 2 metadata files. one has the cluster assignments (above^) and this one here has cell names associated to the clusters
cellNames <- readRDS("/Users/xulab/Documents/Rob/changran analysis/Molecular Diversity and Specialization/annotation.BrainCellAtlas_Saunders_version_2018.04.01.RDS")
cellNames <- filter(cellNames, tissue == "FC")

# outcome = metadata. there are 17 clusters
outcome <- readRDS("/Users/xulab/Documents/Rob/changran analysis/Molecular Diversity and Specialization/F_GRCm38.81.P60Cortex_noRep5_FRONTALonly.cell_cluster_outcomes.RDS")
outcome$rowNames <- row.names(outcome) #the cortex is different than the hippocampus

ref <- data.frame(cluster = factor(c(1:14)),
                  cellType = c(
                    "Neuron",
                    "Neuron",
                    "Neuron",
                    "Neuron",
                    "Neuron",
                    "Neuron",
                    "Neuron",
                    "Astrocyte",
                    "Oligodendrocyte",
                    "Polydendrocyte",
                    "Microglia",
                    "Endothelial Stalk",
                    "Mural",
                    "Endothelial Tip"
                  ))
test <- left_join(outcome, ref, by=join_by(cluster))
# cluster <- readRDS("/Users/xulab/Downloads/F_GRCm38.81.P60Hippocampus.cluster.assign.RDS")
# subcluster <- readRDS("/Users/xulab/Downloads/F_GRCm38.81.P60Hippocampus.subcluster.assign.RDS")


#never use the merge function because it changes the order of the df and fucks up seurat. always use dplyr join
meta <- seu@meta.data
meta$rowNames <- rownames(seu@meta.data)
test <- left_join(meta, test, by=join_by(rowNames))
row.names(test) <- test$rowNames
test$rowNames <- NULL


seu <- AddMetaData(seu, metadata = test)

# filter out low quality cells (anything with a "reason" option in the metadata)
#this is required to form the the 17 clusters discussed in the paper. without this filtering there will be 22 clusters
seu@meta.data$reason <- lapply(seu@meta.data$reason, toString)
seu <- subset(seu, subset = reason == "NA")


#look at tally of clusters
# seu@meta.data %>%
#   group_by(cluster) %>%
#   tally()

#look at tally of subclusters
# seu@meta.data %>%
#   group_by(subcluster) %>%
#   tally()

View(seu@meta.data)
# unique(outcome$cluster)
# sort(unique(seu@meta.data$cluster))

```

data processing from youtube video:


# 1. QC -------
```{r}

# mito_genes <- grep("^mt-", rownames(seu), value = TRUE)
# seu[["percent.mt"]] <- PercentageFeatureSet(seu, features = mito_genes)

seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^mt-")

# View(seu@meta.data)
#expression of one mito gene
# seu[["RNA"]]$counts[24463, ]

#list of mito genes
# grep("^mt-", rownames(seu[["RNA"]]$counts))

#name of one mito gene
# rownames(seu[["RNA"]]$counts)[24463]
Idents(seu) <- "cellType"

VlnPlot(seu, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
FeatureScatter(seu, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") +
  geom_smooth(method = 'lm')


```



# 2. Filtering -----------------

cells were already filtered out                          
```{r}
#this is default
# seu2 <- subset(seu, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 &
#                           percent.mt < 5)

# seu <- subset(seu, subset = nFeature_RNA > 200 & percent.mt < 20)

# View(seu@meta.data)

# seu[["RNA"]]$data <- seu[["RNA"]]$counts

# View(seu@assays[["RNA"]]@cells)

```
                          

# 3. Normalize data ----------
take gene expression values of each cell, divide by total gene expression,
multiply by a scaling factor, then log transform. this normalizes expression 
and allows you to compare expression across cells

```{r}
#the default parameters (when you have no parameters specified) are as follows:
#seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)

seu <- NormalizeData(seu)

# View(seu2@meta.data)
# View(seu@meta.data)
# str(seu)

```



# 4. Identify highly variable features --------------
2000 most highly variable genes are saved
```{r}
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(seu), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seu)
LabelPoints(plot = plot1, points = top10, repel = TRUE)
```


# 5. Scaling -------------
the data will have unwanted sources of variation. technical noise, such as 
batch effect, differences in cell cycle, etc. 
```{r}
# all.genes <- rownames(seu)
seu <- ScaleData(seu, features = rownames(seu))

```


# 6. Perform Linear dimensionality reduction --------------
```{r}
seu <- RunPCA(seu, features = VariableFeatures(object = seu))
VizDimLoadings(seu, dims = 1:2, reduction = "pca")
```


# visualize PCA results
```{r}
print(seu[["pca"]], dims = 1:5, nfeatures = 5)
DimHeatmap(seu, dims = 1, cells = 500, balanced = TRUE)
```


# determine dimensionality of the data.
this is also known as a scree plot
```{r}
ElbowPlot(seu)
```

it appears from the elbow plot that most of the variation in the dataset is captured within the first 15 principal components
# 7. Clustering ------------
```{r}
seu <- FindNeighbors(seu, dims = 1:15)
```


# understanding resolution
```{r}
# this is the default, it has a lot of clusters
# seu <- FindClusters(seu, resolution = c(0.1,0.3, 0.5, 0.7, 1))
seu <- FindClusters(seu, resolution = c(0.03, 0.05, 0.1, 0.2))

# View(seu@meta.data)

DimPlot(seu, group.by = "RNA_snn_res.0.05", label = TRUE)
```



# non-linear dimensionality reduction --------------
```{r}
seu <- RunUMAP(seu, dims = 1:15)
DimPlot(seu, reduction = "umap")
```


```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
seu.markers <- FindAllMarkers(seu, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
seu.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
```



visualize how good your clusters are 
```{r}
seu.markers %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(seu, features = top10$gene) + NoLegend()
```

```{r}
# unique(Idents(seu))
seu2 <- seu

seu <- subset(seu, idents = c("Neuron", "Astrocyte", "Oligodendrocyte", "Microglia"))
Idents(seu) <- factor(seu$cellType, levels = c(
                    "Neuron",
                    "Astrocyte",
                    "Oligodendrocyte",
                    "Microglia"
                  ), ordered = TRUE)
# View(seu@meta.data)
# Idents(seu) <- cellTypeFactor
```


```{r}
features <- c("Ntrk2")
RidgePlot(seu,
          features = features,
          # sort = "decreasing"
          # cols = #colors of groups
          # sort = c("WT 7 week","WT 1 month", "WT 3 month",
          #          "WT 6 month", "WT 8 month","WT 20 month",
          #          "5xFAD 1 month", "5xFAD 3 month", "5xFAD 6 month",
          #          "5xFAD 9 month", "WT 6 month ctx", "5xFAD 6 month ctx"
          #          )
          #change the order depicted
          # group.by = seu@meta.data$group,
          # ncol = 1 #only use if multiple plots are displayed
          ) 
```

```{r}
features <- c("Bdnf")
bdnfCtx <- VlnPlot(seu,
          features = features
          ) +
  ggtitle("Bdnf Expression in Frontal Cortex") +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.8, size = 0.5
  ) 
  # geom_dotplot(binaxis= "y", alpha = 0.01)
  # geom_jitter(color = "gray")
  # scale_fill_manual(values = "black")
bdnfCtx
```


```{r}
features <- c("Ntrk2")
Ntrk2Ctx <- VlnPlot(seu,
          features = features,
          # cols = #colors of groups
          # sort = "increasing"
          #change the order depicted
          # group.by = seu@meta.data$group,
          # ncol = 1 #only use if multiple plots are displayed
          ) +
  ggtitle("Ntrk2 Expression in Frontal Cortex") +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.8, size = 0.5
  )
Ntrk2Ctx
```

```{r}
#cowplot:
# grid <- plot_grid(bdnfCtx, Ntrk2Ct2, labels = c("A", "B"))
# grid

# ggpubr:
# grid <- ggarrange(bdnfCtx, Ntrk2Ctx)
grid <- ggarrange(bdnfCtx, Ntrk2Ctx, bdnfHippo, ntrk2Hippo)
grid
```

ˆ
```{r}
features <- c("Bdnf", "Ntrk2")
Percent_Expressing(seu, features)
Percent_Expressing(hseu, features)
# DotPlot(seu, features = features) + RotatedAxis()
```

```{r}
ggexport(grid, filename="ctxHippoExpression.pdf", width = 14, height = 16)
```


this saves the entire seurat object
```{r}
# saveRDS(seu, file = "../GSE140511 5x all cells /seu")
```


